Estimating Gaze Duration Error from Eye Tracking Data

### 2023-03-01 acm_version/paper.tex
ACM Version of Paper

\documentclass[12pt,a4paper]{article}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{booktabs}
\usepackage[latin2]{inputenc}
\usepackage{t1enc}
\usepackage[mathscr]{eucal}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{graphics}
\numberwithin{equation}{section}
\usepackage{hyperref}
\usepackage[margin=2.9cm]{geometry}
\usepackage[outdir=./images/]{epstopdf}
\epstopdfsetup{outdir=./images/}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\author{
John Hawkins \\ email \href{mailto:john.hawkins@playgroundxyz.com}{john.hawkins@playgroundxyz.com} \\
}

@@

\section{Abstract}

Eye tracking technologies produce a series of timestamped gaze fixation points 
that can be attributed to objects within a subject's field of vision. 
These fixation points are often aggregated in order to generate a measurement
of the gaze duration in Area of Interest studies.
Error is typically measured on the basis of the variation between individual 
gaze fixation points and the intended point of fixation.
These underlying errors in fixation have no inherent transformation for making
a statement about error in gaze duration, even though they are expected to be
related.

In this work we develop an algorithm for estimating the error distribution of 
gaze duration measurement through Monte Carlo simulation using the content 
of an eye tracking calibration log file. 
We provide this algorithm in an open source application
to allow researchers to understand the error in their gaze duration measurements. 
We use this application to conduct experiments on the expected error bounds for 
different duration measurements across a fixed session length for a simulated 
area of interest study on a mobile device. 
The results indicate that error in gaze duration estimation is sensitive to
fixation error beyond a bound that will depend on the size of the area of interest 
and the comparative length of the viewing session.

@@

The source repository for the \textit{gazerr} application also contains a series of 
scripts for generating data sets and running the experiments outlined in the next 
section. Full details for the data generation process are available in the 
source code README file.

@@

We apply the \textit{gazerr} application to investigate the relationship between 
measured gaze duration and expected true duration under a variety of scenarios. 
Each scenario involves a variation in the underlying error indicated by the eye tracking 
calibration file and length of the session in which the user could be looking at 
the area of interest.

For the sake of our experiments we use a simulated device with viewport pixel 
dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium 
rectangle ad unit (MREC) that is located toward the top of the screen. 
We iterate over a range of potential mean error values 
(equivalent to varying the accuracy), determined to by the mean Euclidean distance 
between the measured fixation point and actual fixation points.
For each of these mean error values we generate a synthetic eye tracking calibration 
file. We produce three versions of this calibration data, one in which the error 
is centered around the true target, a second biased version in which the error tends 
toward the upper left of the true position. In the third version the error is again
biased toward the upper left corner of the screen, but it has higher precision, 
meaning that the measurements will cluster closely together with each other around
the mean error position.

We include these biased calibration datasets as they are consistent with both our 
observations of real eye tracking data and what is found in the literature\cite{Holmqvist2022}.
We use this synthetic data creation in order to explore the extent to which different
patterns in error, bias and precision affects gaze duration measurements.

@@

In these experiments we use both a fixed session length and fixed area of interest
size and position, so that we can focus specifically on the impacts of mean
error, bias and precision on the quality of gaze duration measurment.

@@

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
\caption{Measured Gaze Duration Versus Expected Gaze Duration. \\\hspace{\textwidth} 
MREC ad appearing on screen for 1s with 70px Mean Error in Fixation}
\label{fig:measured_vs_expected}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{../results/Fixation_vs_duration_error.png}
\caption{Fixation error versus gaze duration error.}
\label{fig:fixation_vs_duration}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{../results/Error_surface.png}
\caption{Mean gaze duration error over measured error and fixation MAE.}
\label{fig:error_surface}
\end{figure}

@@

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}

>>>>

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%  
%% The original source files were: 
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[MLHCI '23]{}{March 24--26,2023}{Singapore}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/23/05}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
\acmSubmissionID{VP5015}

%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%
%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{booktabs}
\usepackage[latin2]{inputenc}
\usepackage{t1enc}
\usepackage[mathscr]{eucal}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{graphics}
\numberwithin{equation}{section}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{John Hawkins}
\email{john.hawkins@playgroundxyz.com}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Playground XYZ}
  \streetaddress{Level 13/50 Carrington St}
  \city{Sydney}
  \state{NSW}
  \country{Australian}
  \postcode{2000}
}
\renewcommand{\shortauthors}{Hawkins}

\begin{abstract}
Eye tracking technologies produce a series of timestamped gaze fixation points 
that can be attributed to objects within a subject's field of vision. 
These fixation points are often aggregated in order to generate a measurement
of the gaze duration in Area of Interest studies.
Error is typically measured on the basis of the variation between individual 
gaze fixation points and the intended point of fixation.
These underlying errors in fixation have no inherent transformation for making
a statement about error in gaze duration, even though they are expected to be
related.

In this work we develop an algorithm for estimating the error distribution of 
gaze duration measurement through Monte Carlo simulation using the content 
of an eye tracking calibration log file. 
We provide this algorithm in an open source application
to allow researchers to understand the error in their gaze duration measurements. 
We use this application to conduct experiments on the expected error bounds for 
different duration measurements across a fixed session length for a simulated 
area of interest study on a mobile device. 
The results indicate that error in gaze duration estimation is sensitive to
fixation error beyond a bound that will depend on the size of the area of interest 
and the comparative length of the viewing session.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10003260.10003272</concept_id>
       <concept_desc>Information systems~Online advertising</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002951.10003260.10003261.10003267</concept_id>
       <concept_desc>Information systems~Content ranking</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002951.10003260.10003277.10003281</concept_id>
       <concept_desc>Information systems~Traffic analysis</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Online advertising}
\ccsdesc[500]{Information systems~Content ranking}
\ccsdesc[500]{Information systems~Traffic analysis}

\keywords{Eye Tracking, Attention Measurement, Digital Advertising, Error Estimation}

@@

@@

The source repository for the \textit{gazerr} application also contains the complete
set of scripts for re-creating the data sets and running the experiments outlined in the next 
section. Please refer to the README file for the sequence of steps required for replication
of all results.

@@

We apply the \textit{gazerr} application to investigate the relationship between 
measured gaze duration and expected true duration of fixation under a variety of 
simulated scenarios. 
Each scenario involves a variation in the underlying error indicated by the eye tracking 
calibration file and length of the session in which the user could be looking at 
the area of interest.

For the sake of our experiments we use a simulated device, approximately the size of a
small smart-phone, with viewport pixel dimensions of 350 by 627. 
Our area of interest (AOI) is a fixed position 
rectangle that resembles an iAB standard medium rectangle ad unit (MREC). We position the
simulated ad toward the top of the screen. 
We iterate over a range of potential mean error values from 20 to 160 pixels 
(equivalent to varying the accuracy of the eye tracking models). These mean error
values determine the average Euclidean distance 
between a measured fixation point and the true fixation point.

For each of these mean error values we can generate synthetic eye tracking calibration 
files containing a sample of true fixation points and measured fixation points that are
coherent with the simulation parameters. 
We produce three versions of this calibration data to simulate and experiment with
multiple real-world scenarios:

\begin{itemize}
    \item {\texttt{Unbiased}}: Error distributed uniformly around true fixation point. 
    \item {\texttt{Biased}}: Error tending toward upper left of true fixation.
    \item {\texttt{Biased Precise}}: Error tending toward upper left but clustered.
\end{itemize}

These three scenarios are listed above. The unbiased calibration data is
distributed evenly and centered around each of the true target positions in the callibration
file. The biased data imposes a tendency on the error to be pushed toward the upper left 
of the screen away from the true fixation position. The biased precise data is again forced
toward the upper left corner of the screen, but constrained so that measurements cluster
together, meaning that although measurements have the same mean error that are more tightly
clustered with each other around the mean error position.

We include the two biased calibration datasets as they are consistent with both our 
observations of real eye tracking data and what is found in the literature\cite{Holmqvist2022}.
We use this synthetic data creation in order to explore the extent to which different
patterns in error, bias and precision will affects gaze duration measurements.

@@

In these experiments we use both a fixed session length of one thousand milliseconds
and a fixed position for the area of interest. These restrictions allow us to focus 
on the impacts of mean error, bias and precision on the quality of gaze duration measurment.

@@

\begin{figure*}[h!]
\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
\caption{Measured Gaze Duration Versus Expected Gaze Duration. \\\hspace{\textwidth} 
MREC ad appearing on screen for 1s with 70px Mean Error in Fixation}
\label{fig:measured_vs_expected}
\end{figure*}

@@

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.4]{../results/Fixation_vs_duration_error.png}
\caption{Fixation error versus gaze duration error.}
\label{fig:fixation_vs_duration}
\end{figure*}

@@

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.4]{../results/Error_surface.png}
\caption{Mean gaze duration error over measured error and fixation MAE.}
\label{fig:error_surface}
\end{figure*}

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.
